---
title: "Turkey's Education Levels and Exposure to Print Media"
format: pdf
author: "Youngho Kim, Laura Lee-Chu, Sebastian Rodriguez"
abstract: "" 
date: today
date-format: long
editor: visual
toc: TRUE
number-sections: TRUE
bibliography: references.bib
thanks: "Code and data supporting this analysis is available at: [https://github.com/seb646/turkey-education-media](https://github.com/seb646/turkey-education-media)"
---

```{r}
#| messages: false
#| echo: false
#| warning: false
#| output: false
#### Preamble ####
# Title: Turkey's Education Levels and Exposure to Print Media
# Author: Youngho Kim, Laura Lee-Chu, Sebastian Rodriguez
# Email: me@srod.ca
# Date: 27 March 2023
# GitHub: https://github.com/seb646/turkey-education-media

#### Workspace set-up ####
# Define required libraries 
library(lubridate)
library(tidyverse)
library(dplyr)
library(knitr)
library(janitor)
library(scales)
library(RColorBrewer)
library(ggplot2)
library(kableExtra)
library(here)
library(arrow)

# Fetch the education levels data
education_levels <- arrow::read_parquet(here::here("inputs/data/education_levels.parquet"), show_col_types = FALSE)

# Fetch the exposure to print media data
media_exposure <- arrow::read_parquet(here::here("inputs/data/exposure_to_print_media.parquet"), show_col_types = FALSE)
```

## Introduction

## Data

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Table 1: Education Levels"
#| label: table-1
education_levels |>
  # Select relevant columns
  select(
    characteristic, 
    no_education, 
    primary_incomplete, 
    primary_complete,
    secondary_incomplete,
    secondary_complete,
    number_of_people
  ) |>
  
  # Rename column rows
  rename(
    "Background Characteristic" = "characteristic", 
    "No Education" = "no_education", 
    "Primary Incomplete" = "primary_incomplete", 
    "Primary Complete" = "primary_complete", 
    "Secondary Incomplete" = "secondary_incomplete",
    "Secondary Complete+" = "secondary_complete",
    "Number of People" = "number_of_people"
  ) |>
  
  # Init the table
  kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "") |>
  
  # Add headers
  add_header_above(
    c("", "Level of Education" = 5, ""),
    align = "c",
    bold = TRUE
  ) |>
  
  # Style the table 
  kable_styling(font_size = 8, latex_options = "hold_position", full_width = TRUE) |>
  
  # Increase width of first column
  column_spec(1, width = "13em")|>

  # Add section headers
  pack_rows("Women - Age", start_row = 1, end_row = 7) |>
  pack_rows("Women - Residence", start_row = 8, end_row = 9) |>
  pack_rows("Husbands - Age", start_row = 10, end_row = 16) |>
  pack_rows("Husbands - Residence", start_row = 17, end_row = 18)
  
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Table 2: Exposure to Print Media"
#| label: table-2
media_exposure |>
  # Rename column rows
  rename(
    "Background Characteristic" = "characteristic", 
    "Weekly Newspaper Reading" = "women_read_weekly", 
    "Number of Women" = "women_number", 
    "Weekly Newspaper Reading\n" = "husband_read_weekly", 
    "Number of Husbands" = "husband_number"
  ) |>
  
  # Init the kable table
  kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "") |>
  
  # Add headers
  add_header_above(
    c("", "Women" = 2, "Husbands" = 2),
    align = "c",
    bold = TRUE
  ) |>
  
  # Style the table 
  kable_styling(font_size = 8, latex_options = "hold_position", full_width = TRUE) |>
  
  # Increase width of first column
  column_spec(1, width = "13em")|>
  
  # Add section headers
  pack_rows("Age", start_row = 1, end_row = 8) |>
  pack_rows("Residence", start_row = 9, end_row = 10)
  
```

## Discussion

## Inclusive Acknowledgements

**Land Acknowledgement**

We wish to acknowledge the land on which this data was analysed. For thousands of years, it has been the traditional land of the Huron-Wendat, the Seneca, and the Mississaugas of the Credit. Today, this meeting place is still the home to many Indigenous people from across Turtle Island and we are grateful to have the opportunity to work on this land.

This statement was originally prepared in consultation with First Nations House and the Elders Circle for use at the University of Toronto [@cite_land_acknowledgement].\

**Data Acknowledgement**

The data used in this paper was obtained from the Demographic and Health Surveys (DHS), a program of the United States Agency for International Development (USAID). The DHS and USAID gathered this data from the 1998 Turkey Demographic and Health Survey, originally collected by Hacettepe University, Institute of Population Studies and Macro International.

**Resources Acknowledgment**

The primary tool used to analyse data in this paper is R, an open-source statistical programming language [@cite_tool_r]. The paper also uses a number of R packages, including: dplyr [@cite_tool_dplyr], foreign [@cite_tool_foreign], ggplot2 [@cite_tool_ggplot2], here [@cite_tool_here], janitor [@cite_tool_janitor], kableExtra [@cite_tool_kableextra], knitr [@cite_tool_knitr], lubridate [@cite_tool_lubridate], readr [@cite_tool_readr], RColorBrewer [@cite_tool_rcolorbrewer], scales [@cite_tool_scales], and tidyverse [@cite_tool_tidyverse].

## References

<!-- Comment required to show references as a numbered section -->
